---
title: "Pre-midterm Section"
author: "Hikaru Yamagishi"
date: "3/7/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Today in section, we reviewed the bootstrap. First, conceptually; then in R code.

This code contains two ways to go about bootstrapping: We can define a function or use a `for` loop. The first may be more helpful for the near future.


# Bootstrapping: Concepts

* Plug-in principle
* ECDF
* "Pull yourself up by your bootstraps"

Remember that we are trying to characterize the uncertainty of our estimator.


# Bootstrapping: 4 steps

First, let us walk through the steps of bootstrapping. 

We will talk through each step and figure out the code for **one** resampling iteration first.

We will start with one variable. Remember Pset 2? What was the setting then? Remember to keep population and sample straight.

## Data

Now let's assume that this example data `df_eg1` is our sample data ($n=100$):

```{r}
# Let's say this is our sample data
df_eg1 <- sample(x = c(56:59), size = 100, replace = TRUE)
head(df_eg1)
```

## 4 steps

1. From your sample, take a *with replacement* sample of size $n$.

```{r}
# Define the number of samples you want to draw
n_data <- length(df_eg1)

# Sample
samp <- sample(x = df_eg1, size = n_data, replace = TRUE)
samp
```

2. Calculate what your estimate would be using the bootstrap sample from step 1.

If the test statistic of interest was the mean, we would run:

```{r}
mean(samp)
```

3. Repeat step 1 and step 2 many times.

We now have a collection of bootstrap estimates.

4. Calculate the standard deviation of the bootstrap distribution of our estimator.

We take the bootstrap estimates obtained in step 3 and run the `sd()` function.


# Example 2: Full example

Now suppose there are multiple variables, $X$, $Y$, and $Z$.

Suppose that our sample statistic of interest is the sum of the mean of $X$, mean of $Y$, and mean of $Z$. Now we will try the whole bootstrapping procedure, including the iteration part.


## Data

Let's use this toy data.

```{r}
set.seed(503)

data_sample <- data.frame(
  x = rnorm(n = 1000, mean = 3, sd = 1),
  y = rnorm(n = 1000, mean = 4, sd = 2),
  z = rnorm(n = 1000, mean = 6, sd = 3)
)
```

## Define the function

Let's define our function `estimator`:

```{r}
estimator <- function(sample) {mean(sample$x) + mean(sample$y) + mean(sample$z)}
estimator(sample = data_sample)
```

## Bootstrapping using a user-defined function

1. From your sample, take a *with replacement* sample of size $n$. This time, let's define this as a function. Notice how we sample different variables. *discussion*

```{r}
draw_bootstrap_sample <- function() {
  data_sample[sample(nrow(data_sample), replace = TRUE), ]
}
```

Addendum: `samp.int` is a bare interface version of `sample`. *discussion*

2. Calculate what your estimate would be using the bootstrap sample from step 1.
3. Repeat step 1 and step 2 many times.

```{r}
n_iter <- 1000
bootstrap_est <- replicate(n_iter, estimator(draw_bootstrap_sample()))
```

We now have a collection of bootstrap estimates.

4. Calculate the standard deviation of the bootstrap distribution of our estimator (standard error)

```{r}
est_se_est <- sd(bootstrap_est)
est_se_est
```


# Example using `for` loop

Suppose we have a function that will give us three values for each of the three variables.

## Define function

Don't worry about what this is, we just want to illustrate how we may do this with three outputs (as discussed in class).

```{r}
estimator2 <- function(sample) {
  c(mean(sample$x), mean(sample$y)/mean(sample$x), mean(sample$z)/mean(sample$x))
}
estimator2(sample = data_sample)
```

## Using a `for` loop

```{r}
# Number of sample data rows
nrow_samp <- nrow(data_sample)

# Number of iterations
n_iter = 1000

# Store the three coefficients from each iteration in a row of this matrix.
boot <- matrix(NA, nrow = n_iter, ncol = 3)

# For loop 
for(i in 1:n_iter){
  samp_data <- data_sample[sample(nrow_samp, replace = T),]
  boot[i,] <- estimator2(samp_data)
}

# Take sd() of each column to return a bootstrapped SE on each estimate.
apply(boot, MARGIN = 2, FUN = sd)
```


